---
title: "p8105_homework_1"
output: github_document
author: "Yuchen Zhang"
date: 2022-09-24
---

```{r setup, message=FALSE}
library(tidyverse)
```

# Problem 1

Load the penguins database.
```{r } 
data("penguins", package = "palmerpenguins")
```
The dataset has the following variables.
```{r echo = FALSE}
colnames(penguins)
```
This dataset includes information of Adelie, Chinstrap and Gentoo three species of penguins, their location of Biscoe, Dream or Torgersen. Their bill length ,bill depth, flipper length, body mass,sex and year.

The dataset have `r ncol(penguins)` columns and `r nrow(penguins)` rows. the mean flipper length is `r mean(na.omit(penguins$flipper_length_mm))`.

A scatterplot of flipper_length_mm (y) vs bill_length_mm (x) and export to project directory.
```{r scatterplot, warning=FALSE} 
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + labs(x = "Bill Length(mm)", y = "flipper length(mm)") +
  ggtitle("Scatter plot of Bill Length and flipper length")

ggsave("scatter_plot.pdf", height = 4, width = 6)
```

# Problem 2

Create a data frame
```{r data frame} 

  df = tibble(
             sample = rnorm(10),
             vec_logical = sample > 0,
             vec_character = c('a','b','c','d','e','f','g','h','i','j'),
             vec_factor = factor(c("x","x","y","z","x","y","z","z","y","y"))
  )
```
The mean of each variable in dataframe.
```{r, warning=FALSE}  
  print(mean(pull(df, sample)))
  print(mean(pull(df, vec_logical)))
  print(mean(pull(df, vec_character)))
  print(mean(pull(df, vec_factor)))
```
The mean function mean() works on the numeric vector and logical vector. But, it cannot work on character vector and factor vector.

Convert variables
```{r, warning=FALSE, results = 'hide'} 
  as.numeric(pull(df, vec_logical))
  as.numeric(pull(df, vec_character))
  as.numeric(pull(df, vec_factor))
```
as.numeric can convert logical vector to numeric values 1 or 0 and factor vector to numeric values according to their level. However, character vector is converted to NA.

This explains what happened when I take the mean. mean() can convert logic value and numeric value to numbers to operate. mean() cannot convert factor value directly to numeric value, so it cannot return mean for factor vector.
